### 2021-12-15 corrected_figures.Rmd
results corrected

---
title: "Corrected Figures"
author: "Kevin Kiley"
date: "12/15/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggdag)
library(ggrepel)
library(here)
library(tidyverse)
'%!in%' <- function(x,y)!('%in%'(x,y))
```

```{r, message = FALSE, warn = FALSE}
load(here("correction", "results_df.Rdata"))
load(here("data", "attitude_vars.Rdata"))

label_vars <- res.df %>%
  filter(var %in% c("intlblks", "confed", "selfhunt", "socbar")) %>%
  left_join(attitude_vars)

#create a list of variables that display updating
aum_vars <- res.df %>% filter(aum_pref == 1)
aum_vars <- aum_vars$var

```

```{r loadfiguredata}
load(here("correction", "figure_data.Rdata"))
```

```{r loadcoarse}
#load(here("correction", "coarse_vars.Rdata"))

```

```{r phihist, message=FALSE, fig.width=6.8, fig.cap="Distribution of phi estimates and probabilities that items show evidence of active updating."}

res.df %>%
  filter(grepl("_c", var)==FALSE) %>%
  select(var, p, prob_aum) %>%
  gather(key = "key", value = "value", -var) %>%
  mutate(key = recode(key, "p"="Phi estimates (Updating)", "prob_aum"="Probability of some active updating")) %>%
  ggplot(aes(x = value, y=..count../sum(..count..))) +
  geom_histogram(fill = "gray", color = "black", bins = 15) + 
  geom_vline(xintercept = .5, color = "black", linetype = 2) + 
  theme_classic() + 
  facet_wrap(~key, scales = "free_x") +
  expand_limits(x = 1) +
  labs(x = "", y = "Proportion of items") + 
  theme(axis.title=element_text(size=9))
```

```{r agegroupcomparison, message=FALSE, fig.width=6.8, fig.height=8, fig.cap="Comparison of phi estimates for individuals over and equal to or less than 30 years old, with 95 percent confidence intervals."}

load(here("correction", "age_results.Rdata"))

bind_rows(age.results) %>%
  mutate(ci.lower = estimate - 1.96 * std.error,
         ci.upper = estimate + 1.96 * std.error) %>%
  left_join(attitude_vars) %>%
  mutate(age = recode(age, "old"="Over 30", "young"="30 & under")) %>%
  ggplot(aes(x = reorder(var.name, estimate), y = estimate, color = age)) + 
  geom_hline(yintercept = .5, color = "black", linetype = 2) + 
  geom_pointrange(aes(ymin = ci.lower, ymax = ci.upper)) + 
  coord_flip() + 
  theme_minimal() + 
  scale_color_manual(values = c("black", "gray")) + 
  labs(x = "", y = "Phi estimate (Updating)", color = "") +
  theme(axis.text=element_text(size=7),
        axis.title=element_text(size=9),
        legend.position = "bottom")

```

```{r phibetascatter, message=FALSE, fig.width=6.8, fig.height = 6.0, fig.cap="Distribution of beta and phi estimates for GSS items, by whether model prefers phi = .5 constraint."}
res.df %>%
  filter(grepl("_c", var)==FALSE) %>%
  mutate(aum = ifelse(var %in% aum_vars, "\nPersistent\nChange\n", "\nNo\nPersistent\nChange\n")) %>%
  ggplot(aes(x = p, y = b1)) + 
  geom_point(aes(fill = aum, shape = aum), alpha = .7) + 
  scale_shape_manual(values = c(21, 22)) + 
  scale_fill_manual(values = c("gray", "black")) + 
  geom_text_repel(data = label_vars, aes(label = var.name), size = 3) + 
  theme_minimal() +
  labs(x = "Phi estimate (Updating)", y = "Beta estimate (Consistency)", fill = "", shape = "") + 
  theme(axis.title=element_text(size=9))
```

```{r summary, message=FALSE, fig.width=6.8, fig.height=9.0, fig.cap="Summary of phi estimates for all items, by topical group. Numbers in parentheses indicate the number of items in each topical group."}
load(here("correction", "summary_plot_data.Rdata"))

cat_counts <- attitude_vars %>% group_by(small.cat) %>% summarise(n = n())

plot.data %>%
  left_join(cat_counts) %>%
    mutate(small.cat = ifelse(small.cat == "Sex, sexuality & abortion",
                            "Sex, sexuality &\nabortion", small.cat),
           small.cat = ifelse(small.cat == "Guns, laws, crime & police",
                            "Guns, laws, crime &\npolice", small.cat),
          small.cat = ifelse(small.cat == "Confidence in leadership",
                            "Confidence in\nleadership", small.cat),
          small.cat = ifelse(small.cat == "Relig. identity & beliefs",
                            "Relig. identity &\nbeliefs", small.cat),
          var.name = ifelse(var.name == "Executive branch", "Exe. branch", 
                                   var.name)) %>%
  mutate(small.cat = paste(small.cat, " (", n, ")", sep = "")) %>%
  ggplot(aes(x = reorder(small.cat, median))) + 
  geom_pointrange(aes(y = median, ymin = q25, ymax = q75)) + 
  geom_point(aes(y = min), shape = 21) + 
  geom_point(aes(y = max),shape = 21) + 
  geom_text(aes(y = max, label = var.name), size = 2, position = position_nudge(y = .07)) + 
  coord_flip() + 
  theme_minimal() + 
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank(),
    strip.placement = "outside",
    legend.position = "none",
    legend.title = element_blank(),
    axis.text=element_text(size=7),
    legend.text=element_text(size=9)
    )  +
  expand_limits(y = .9) +
  labs(x = "", y = "Phi estimate (Updating)")

```

```{r polvariables, message=FALSE, fig.width=7, fig.height=8, fig.cap="Phi (updating) and beta (consistency) estimates for 43 political attitudes and identities."}
left_join(res.df, attitude_vars) %>%
  filter(small.cat %in% c("Politics & government", "Public spending",
                          "Sex, sexuality & abortion") |
           var %in% c("cappun", "divlaw", "gunlaw", "grass", "courts",
                      "letin1a")) %>%
  mutate(small.cat = ifelse(small.cat %in% c("Race & immigration", 
                                             "Guns, laws, crime & police") | 
                              var %in% c("prayer", "uswary"), 
         "Misc. policy" , small.cat)) %>%
  filter(var %!in% c("abscale6", "absingle", "teensex", "xmarsex",
                     "homosex", "premarsx")) %>%
  mutate(var.name = gsub("Government", "Gov't", var.name),
         var.name = gsub("Abortion", "Abort", var.name),
         small.cat = ifelse(small.cat == "Sex, sexuality & abortion",
                            "Abortion & sex", small.cat)) %>%
  mutate(small.cat = factor(small.cat, c("Politics & government", "Public spending",
                                        "Abortion & sex", "Misc. policy"))) %>%
  ggplot(aes(x = p, y = b1)) + 
  geom_vline(xintercept = .5, linetype = 2, color = "gray") + 
  geom_point(aes(fill = small.cat), shape = 21) + 
  geom_text_repel(aes(label = var.name), size = 1.75) + 
  theme_bw() + 
  labs(x = "Phi estimate (Updating)", y = "Beta estimate (Consistency)", fill = "", shape = "") + 
  theme(axis.title=element_text(size=9)) +
  facet_wrap(~small.cat) + 
  theme(legend.position = "none")

```

```{r religphi, message=FALSE, fig.width=6.8, fig.height=8.5, fig.cap="Phi estimates for items about religious activity and beliefs, social life, subjective SES, and suicide."}
clean.pestimates %>%
  filter(grepl("_c", var)==FALSE) %>%
  filter(small.cat %in% c("Relig. Activity", "Relig. identity & beliefs",
                          "Social life", "Subjective SES", "Suicide")) %>%
  ggplot(aes(x = reorder(var.name, estimate), y = estimate, color = as.factor(stable))) + 
  geom_hline(yintercept = .5, linetype = 2, color = "gray") + 
  geom_pointrange(aes(ymin = ci.lower, ymax = ci.upper)) + 
  geom_point() + 
  coord_flip() + 
  theme_minimal() + 
  facet_grid(small.cat ~ ., scales = "free_y", space = "free", switch = "both") + 
  scale_color_manual(values = c("black", "darkgray")) + 
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank(),
    strip.text.y =element_text(size=8),
    strip.placement = "outside",
    legend.position = "none",
    legend.title = element_blank(),
    axis.text=element_text(size=7),
    legend.text=element_text(size=7)
  )  +
  expand_limits(y = .9) +
  labs(x = "", y = "Phi estimate (Updating)")

```

```{r poliphi, message=FALSE, fig.width=6.8, fig.height=8.5, fig.cap="Phi estimates for items about guns, law, crime and policing; politics and government; and public spending. The item 'Police can hit citizens', which has phi = .43, has been removed for ease of viewing"}

clean.pestimates %>%
  filter(grepl("_c", var)==FALSE) %>%
  filter(small.cat %in% c("Guns, laws, crime & police", "Politics & government",
                          "Public spending")) %>%
  filter(var != "polhitok") %>%
  ggplot(aes(x = reorder(var.name, estimate), y = estimate, color = as.factor(stable))) + 
  geom_hline(yintercept = .5, linetype = 2, color = "gray") + 
  geom_pointrange(aes(ymin = ci.lower, ymax = ci.upper)) + 
  geom_point() + 
  coord_flip() + 
  theme_minimal() + 
  facet_grid(small.cat ~ ., scales = "free_y", space = "free", switch = "both") + 
  scale_color_manual(values = c("black", "darkgray")) + 
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank(),
        strip.text.y =element_text(size=8),
    strip.placement = "outside",
    legend.position = "none",
    legend.title = element_blank(),
    axis.text=element_text(size=7),
    legend.text=element_text(size=7)
  )  +
  expand_limits(y = .9) +
  labs(x = "", y = "Phi estimate (Updating)")

```

```{r civlibphi, message=FALSE, fig.width=6.8, fig.height=8.5, fig.cap="Phi estimates for items about civil liberties, confidence in leadership, health, morale, and social trust."}
clean.pestimates %>%
  filter(grepl("_c", var)==FALSE) %>%
  filter(small.cat %in% c("Civil liberties", "Trust", "Confidence in leadership",
                          "Health & morale")) %>%
  ggplot(aes(x = reorder(var.name, estimate), y = estimate, color = as.factor(stable))) + 
  geom_hline(yintercept = .5, linetype = 2, color = "gray") + 
  geom_pointrange(aes(ymin = ci.lower, ymax = ci.upper)) + 
  geom_point() + 
  coord_flip() + 
  theme_minimal() + 
  facet_grid(small.cat ~ ., scales = "free_y", space = "free", switch = "both") + 
  scale_color_manual(values = c("black", "darkgray")) + 
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank(),
            strip.text.y =element_text(size=8),
    strip.placement = "outside",
    legend.position = "none",
    legend.title = element_blank(),
    axis.text=element_text(size=7),
    legend.text=element_text(size=7)
  )  +
  expand_limits(y = .9) +
  labs(x = "", y = "Phi estimate (Updating)")
```

```{r racgenphi, message=FALSE, fig.width=6.8, fig.height=8.5, fig.cap="Phi estimates for items about race, gender, sex, sexuality, and abortion."}
clean.pestimates %>%
  filter(grepl("_c", var)==FALSE) %>%
  filter(small.cat %in% c("Gender & family", "Race & immigration", "Sex, sexuality & abortion")) %>%
  ggplot(aes(x = reorder(var.name, estimate), y = estimate, color = as.factor(stable))) + 
  geom_hline(yintercept = .5, linetype = 2, color = "gray") + 
  geom_pointrange(aes(ymin = ci.lower, ymax = ci.upper)) + 
  geom_point() + 
  coord_flip() + 
  theme_minimal() + 
  facet_grid(small.cat ~ ., scales = "free_y", space = "free", switch = "both") + 
  scale_color_manual(values = c("black", "darkgray")) + 
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank(),
                strip.text.y =element_text(size=8),
    strip.placement = "outside",
    legend.position = "none",
    legend.title = element_blank(),
    axis.text=element_text(size=7),
    legend.text=element_text(size=7)
  )  +
  expand_limits(y = .9) +
  labs(x = "", y = "Phi estimate (Updating)")

```